{\pkg}: A Python Package for Integrating Computational Models Across Languages and Scales

### 2019-01-05 papers/lang2018/ms.tex
Updated times and speedup for root/shoot example and added response to reviewer 3

We present an Open Source Python package, {\pkg}, for creating and running integration networks by connecting existing computational models written in different programming languages. {\pkg} was developed as part of the Crops \emph{in Silico} \citep{Marshall-Colon2017} initiative to build a complete crop \emph{in silico} from the level of the genes to the level of the field. {\pkg} is available on Github\footnote{{\tt https://github.com/cropsinsilico/cis\_interface}} with full documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/}} and can be installed from PyPI\footnote{{\tt https://pypi.org/project/cis\_interface/}} via {\tt pip install {\pkg}} or conda-forge\footnote{{\tt https://github.com/conda-forge/cis\_interface-feedstock}} via {\tt conda -c conda-forge install {\pkg}}. Although developed for the purpose of integrating plant biology models, {\pkg} can be applied to any situation requiring coordination between models in the supported languages. \S\ref{S:background} provides background information on existing efforts for cross-language model integration, \S\ref{S:methods} describes the methods used by the {\pkg} package to integrate models, \S\ref{S:results} presents several tests demonstrating the performance of message passing between integrated models, and \S\ref{S:discuss} summarizes the features of the {\pkg} package, describes a few use cases, and outlines areas of ongoing and future development.

@@

{\pkg} is an example of a framework that overcomes these issues by exposing simple and easily accessible interfaces in the language of the model (See \S\ref{SSS:interface}) that permit messages to be passed between the model processes as they run in parallel (See \S\ref{SS:communication}). As a result, models writers only need knowledge of the language in which their model is written.

@@

{\pkg} is built upon several well-established open-source software packages(see below) along with new tools developed explicitly for {\pkg} including utilities for running and monitoring models written in other languages from Python, dynamically creating and managing communication networks, data type conversions between languages, and asynchronous message passing with variable underlying communication mechanisms. {\pkg} currently support models written in Python, Matlab, C, and C++ with additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}. Support for additional languages is planned for future development (See \S\ref{SS:langsupport}).

@@

Model YAMLs include information about the location of the 
model source code, the language the model is written in, how the model should 
be run, and any input or output variables including their data type and units. 
%
%For example, Listings \ref{lst:yml:root} and \ref{lst:yml:shoot} from \A\ref{A:yml} are the model YAMLs for the example root and shoot growth models from \S\ref{S:intro}. These model YAMLs could be used in any integration without being modified because the model inputs and outputs will not change.
%
%\rootyml
%
%\shootyml
%
Integration networks are specified by declaring the connections 
between models and connections are declared by pairing an output variable from 
one model with the input variable of another model. The {\pkg} 
CLI sets up the necessary communication mechanisms to 
then direct data from one model to the next in the specified pattern. Models 
can have as many input and/or output variables as is desired and connections between models are specified by references to the 
input/output variables associated with each model. In addition, input and output variables can also be connected to files.
%
%For example, in the connection YAML for integrating the root and shoot models (Listing \ref{lst:yml:int} from \A\ref{A:yml}), the first two connections link the {\tt root\_input} and {\tt shoot\_input} input channels from the root and shoot model YAML's to input files. The third connection links the {\tt next\_root\_mass} output variable from the root model YAML (Listing \ref{lst:yml:root}) to the {\tt next\_root\_mass} input variable from the shoot model YAML (Listing \ref{lst:yml:shoot}). The forth connection links the {\tt next\_shoot\_mass} to an output file.
%
%\intyml
%
This format allows users the 
flexibility to create complex integration networks and test models in isolation before running the entire integration network.

%---------------------------------------------------------------------------------
% MODEL EXECUTION
\subsubsection{Model Execution}\label{SS:model_execution}
%
{\pkg} launches each model in an integration on its own process, allowing models to complete independent operations in parallel and complete tasks more quickly. For example, the root and shoot growth models from \S\ref{S:intro} require 2.42\,s and 4.06\,s respectively to run for 100 time steps in isolation. If these two models were manually integrated in serial via the command line, the integration would require a total of 6.48\,s. However, the same integration requires only 4.52\,s using {\pkg}, a speed-up of 1.43. The exact speedup provided by parallel integration using {\pkg} depends on how independent the models are and how equally distributed the work is between the models. If the models are fully independent, the speedup is limited by the time required to execute the slowest model. In the example, this would be the shoot growth model which required almost twice the time to execute as the root model. If the models are dependent on one-another, the models will have to wait for messages to be passed and the speedup will not be as great (see \S\ref{SS:speedup} for additional discussion of parallelism and the speedups achievable using {\pkg}). \todo{Fix these numbers for most recent version of the code}.

While every model is executed on a new process, how the model is handled depends 
on the language it is written in. {\pkg} has a dedicated driver for each of the supported languages with utilities that allow {\pkg} to launch and monitor executables written in that language from Python. Models written in 
interpreted languages (Python and Matlab) are executed on the command line 
with the interpreter. In the case of Matlab, where a significant amount of time is required to start 
the Matlab interpreter (See \S\ref{SS:results_language}), Matlab shared engines are used 
to execute Matlab models. Matlab shared engines are Matlab instances that other processes can 
submit Matlab code to for execution. While shared engines also require the same amount of time 
to start as a standard instance of the Matlab interpreter, they can be started in advance and then 
reused.

For the compiled languages (C and C++) there are a few options. The user can 
compile the model themselves, provided they link against the appropriate 
{\pkg} header library. Alternatively, users can provide the location of 
the model source code and let one of several {\pkg} drivers handle the compilation, including 
linking against the appropriate {\pkg} interface library. {\pkg} 
also has support for compiling models using Make \citep{Stallman2004} and 
CMake \citep{Martin2006} for 
models that already have a Makefile or CMakeLists.txt. To use these tools to compile a model, 
lines are added to the recipe in order to allow linking against {\pkg}.

@@

The algorithms for a generic model driver and the master thread can be found in Algorithms \ref{alg:model} \& \ref{alg:master} respectively in \A\ref{A:algorithms}.

@@

%---------------------------------------------------------------------------------
% ZEROMQ
\subsubsection{ZeroMQ}\label{SSS:zmq}
%
The preferred communication mechanism used by {\pkg} are ZeroMQ sockets 
\citep[ZMQ][]{Akgul2013}. ZMQ 
provides broker-less communication via a number of protocols and patterns with 
bindings in a wide variety of languages that can be installed on Posix and Windows 
operating systems. ZMQ was adopted by {\pkg} in order to allow support 
on Windows and for future target languages (See \S\ref{SS:future}) that could not be 
accomplished using IPC queues. In addition, while ZMQ allows 
interprocess communication via IPC, ZMQ also 
supports protocols for distributed communication via an Internet Protocol (IP) 
network. While {\pkg} does not currently support using these protocols 
for distributed integration networks, this is a future development plan that has been prepared for.

In addition to using the default ZMQ libraries, {\pkg} also includes supporting routines for allowing received messages to be confirmed. Because ZMQ is broker-less, a socket has no way of knowing if the message it sent was successfully received. This lack of confirmation makes it harder to determine if there is was error somewhere along the network. {\pkg} overcomes this by generating two ZMQ sockets for every connection: one for sending the messages and one for confirming them. Each time a message is received, the receiving socket confirmation thread will send a confirmation including a unique ID for the received message and then wait for a reply to that confirmation. The sending socket confirmation thread will continuously check the confirmation sockets for messages indicating that sent messages were received. Once a confirmation message is received, it will send a reply to the receiving confirmation socket and record the ID of the message that was confirmed. This handshake operation ensures that all messages are accounted for so {\pkg} knows if a message is lost and where it was lost.

@@

%---------------------------------------------------------------------------------
% INTERFACE
\subsubsection{Interface}\label{SSS:interface}
%
{\pkg} provides interface functions/classes for communication that are 
written in each of the supported languages. Language specific interfaces 
allow users to program in the 
language(s) with which they are already familiar. The Python interface provides 
communication classes for sending and receiving messages. The Matlab 
interface provides a simple wrapper class for the Python class, that exposes 
the appropriate methods and handles conversion between Python and Matlab 
data types. The C interface provides structures and functions for accessing 
communication channels and sending/receiving messages. The C++ interface 
provides classes that wrap the C structures with functions called as methods.

In addition to basic input and output, each interface also provides access 
to more complex data types and communication patterns. Pseudocode describing the asynchronous communication classes and functions underlying the interfaces is provided by Algorithms \ref{alg:async_send} \& \ref{alg:async_recv} in \A\ref{A:algorithms}.

%---------------------------------------------------------------------------------
% SERIALIZATION
%
\subsection{Transformation}\label{SS:transformation}
%
%\subsubsection{Data Formats}\label{SSS:dataformats}
%
Messages are passed as raw bytes. In order to understand the messages begin passed, parallel processes that communicate must agree upon the format used to do so. Without community standards, different models will often use very different data formats for their input and output. Differences between data formats can include, but are not limited to, type, precision, fields, or units. While some data formats are self-descriptive and include these types of information as meta data, this is not true of all data formats. To combat this, {\pkg} requires models to explicitly specify the format of input and output expected by a model in the model YAML. {\pkg} can then handle a number of conversions between models without prompting as well as serialization/deserialization to the correct type in each of the supported languages. Data formats currently supported by {\pkg} include:
%
\begin{itemize}
	\item Scalars (e.g. integers, decimals)
	\item Arrays
	\item Text-encoded tables (e.g. CSV or tab-delimited)
	\item Pandas data frames \citep{pandas}
	\item 3D geometry structures (PLY \citep{ply} \& OBJ \citep{obj})
\end{itemize}

@@

%---------------------------------------------------------------------------------
%---------------------------------------------------------------------------------
% WORKED EXAMPLE
\section{Worked Example}\label{S:example}
%
In order to illustrate how {\pkg} is intended to be used by model writers, the following walks through the integration of the two examples models from \S\ref{S:intro}. All of the source code and YAML files discussed in this section are available in the {\pkg} GitHub repository and will be included with future releases on PyPI and conda-forge. This example assumes that the user starts with the following existing model source code:
%
\rootsrc
%
\shootsrc
%
The root source code in Listing \ref{lst:src:root} is a standalone C header library containing a single function {\tt calc\_root\_mass} that calculates and returns the root mass at the next time step (a double precision floating point number) from input of the relative root growth rate ({\tt r\_r}, double), the time step ({\tt dt}, double), and the root mass as the current time step ({\tt R\_t}, double).
%
The shoot source code in Listing \ref{lst:src:shoot} is a standalone Python module containing a single function {\tt calc\_shoot\_mass} that calculates and returns the shoot mass at the next time step from input of the relative shoot growth rate ({\tt r\_s}), the time step ({\tt dt}, double), the shoot mass at the current time step ({\tt S\_t}), the root mass at the current time step ({\tt R\_t}), and the root mass at the next time step ({\tt R\_tp1}).

@@

For the example presented above, the speedup offered by {\pkg} through parallelism can be parameterized in terms of the durations of calculations for each model (designated model A \& B for brevity) performed in the different phases of the process (See Figure \ref{fig:empty_parallel}). 

@@

%---------------------------------------------------------------------------------
% USE CASES AND LIMITATION
%
\subsection{Advantages \& Limitations}\label{SS:uses}
%
\subsubsection{Advantages}\label{SS:advantages}
%
{\pkg} offers several advantages over manual integrations or other existing tools.
%
\begin{description}
	\item[Dynamic Communication Network] While there are many existing tools for communication that have implementations in multiple languages \citep[some of which {\pkg} uses, e.g.][ZMQ, RabbitMQ, OpenMPI]{ZMQ, RMQ, Gabriel2004}, these tools require that the communication pattern be well established such that it can be hard coded into the model (or model wrapper) itself (e.g. using specific ports or queues). The use of hard coded values greatly restricts the reusability of a model. For every integration a model is used in, the hard coded values for every connection must be updated between the models on either end so that they match. {\pkg} is unique in that it dynamically establishes and manages the networks of communication resources (e.g. queues, channels, sockets) that are required for integrations to run so that models do not need to have these hard coded value and can be reused in multiple integrations, unmodified. This process includes creating new resources, connecting to existing resources, passing the appropriate resource addresses to the model environment, error handling/propagation, and resource cleanup. 
	\item[Hands-off Parallelism \& Asynchronous Communication] In addition to inter-language communication, {\pkg} users benefit from speedups due to parallelism and concurrency without any additional work. From the specified model connections, {\pkg} manages the parallel execution of models and asynchronous message passing under-the-hood so that the user doesn't have to think about it. Existing tools for asynchronous communication between unmanaged parallel tasks \citep[e.g. Jack2,][]{Magoules2018} requires explicit definition of the communication pattern in advance, prohibiting dynamic communication networks as discussed above. Work/data flow tools like Parsl \citep{babuji18} allow dynamic execution of processes along a dependency chain in parallel, but usually require addition code, are limited to passing data as files or data object native to the tool's implementation language, and are ill suited to models with persistent internal states or intermediate outputs. {\pkg} is unique in that it offers users asynchronous communication and parallelism that is flexible enough to support any communication pattern, implicitly handles conversion between native data types in the supported languages, and does not require any additional time investment on the part of the user.
	\item[No New Language] Unlike DSLs or language specific frameworks, {\pkg} doesn't require a model to be written in one specific language. In addition to expanding the possible number of model combinations, this eases the burden on model writers for whom learning an additional programming language or DSL might be a significant obstacle. {\pkg} is designed such that model writers only need to be familiar with the language that the model is written in to being with. Although {\pkg} is a Python package, users do not even need to be familiar with the Python language and can interact with {\pkg} solely through the command line. While users do need to be able to compose YAML specification files in addition to their model code, these files have a very limited syntax and will be GUI composable in future releases.
\end{description}

\subsubsection{Limitations}\label{SS:limitations}
%
While {\pkg} is a powerful tool for integrating scientific models and opening up new avenues for collaboration, it is not applicable to every scenario.
%
\begin{description}
	\item[Single Language Integrations] For models written in the same language, it is usually much more computationally efficient to either integrate the models directly (i.e. by calling one from the other) or use a language specific tools for parallel execution and passing data \citep[e.g. Parsl in Python][]{babuji18} than to use {\pkg}. This performance advantage will be particularly pronounced for models that finish quickly or tightly coupled integrations where models make a large number of exchanges in rapid succession (e.g. a set of coupled ODEs). Although the constituent models will still benefit from integration via {\pkg} in regards to reusability, {\pkg} may not be the best choice in such cases if execution time is a concern.
	\item[Easily Translatable Models] For relatively simple models ($<100$ lines of code), it may be more efficient for a scientist to translate the model into the language of the model(s) it will be integrated with. {\pkg} is still useful as a tool more making the model accessible and more likely to be reused, but the performance advantage of running codes in the same language may be of greater importance in certain situations as discussed above.
	\item[Unsupported Languages] While {\pkg} currently supports models written in Python, Matlab, C, \& C++, models are written in many other languages. There are plans to expand support for {\pkg} to other languages (see \S\ref{SS:langsupport}), but in the meantime {\pkg} will not be particularly useful for models that are not written in one of the languages already supported.
\end{description}

%---------------------------------------------------------------------------------
% FUTURE WORK
\subsection{Current/Future Improvements}\label{SS:future}
%
{\pkg} is being actively developed to expand the number of models 
that can be used in integration networks and the complexity of integration 
networks that can be executed. Several improvements are already in progress/planned for {\pkg}.
\todo{expand?}

% LANGUAGE SUPPORT
\subsubsection{Language Support}\label{SS:langsupport}
%
The biggest barrier to running new models is language support. Plant models 
are written in many different languages. While the current language support 
covers many of the most popular languages among plant biologists, 
additional languages must be added to unlock the full set of potential integrations. 
Surveys of the plant modeling community have helped identify the core languages 
that models have been written in. Based on these results, we plan to expand 
{\pkg} to support models written in R, Fortran, Java, and the SBML \citep{Hucka2003} DSL.

@@

% DISTRIBUTED SYSTEMS
\subsubsection{Distributed Systems}\label{SS:distributed}
%
High performance computing (HPC) and cloud compute resources are powerful 
tools in the current computing ecosystem that could be used for running 
complex integration networks. To this end, we plan to expand {\pkg} 
support for running integration networks on distributed compute resources. 
{\pkg} already uses communication tools that can be adapted for use 
in a distributed pattern. We will leverage tools like the libsubmit package from the 
Parsl project \citep{babuji18} for 
automating the submission process to HPC and compute resources. In addition, 
we will add tools for running models as a service and using RMQ to 
permit access to these models within integration networks.

% CONTROL FLOW
\subsubsection{Control Flow}\label{SS:control}
%
Currently, modelers must explicitly specify how a model should process input 
within the model code including things like which input variables are static 
and received once versus which variables change and should be looped over. 
We plan on adding options to {\pkg} for dynamically generating model code 
based a user provided function call and list of static/variable input channels. 
This will allow better reusability of model code such as during parameter studies.
Control flow will be part of the {\pkg} 1.0 release.

>>>>

We present an Open Source Python package, {\pkg}, for creating and running integration networks by connecting existing computational models written in different programming languages. {\pkg} was developed as part of the Crops \emph{in Silico} \citep{Marshall-Colon2017} initiative to build a complete crop \emph{in silico} from the level of the genes to the level of the field. {\pkg} is available on Github\footnote{{\tt https://github.com/cropsinsilico/cis\_interface}} with full documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/}} and can be installed from PyPI\footnote{{\tt https://pypi.org/project/cis\_interface/}} via {\tt pip install {\pkg}} or conda-forge\footnote{{\tt https://github.com/conda-forge/cis\_interface-feedstock}} via {\tt conda -c conda-forge install {\pkg}}. Although developed for the purpose of integrating plant biology models, {\pkg} can be applied to any situation requiring coordination between models in the supported languages. \S\ref{S:background} provides background information on existing efforts for cross-language model integration, \S\ref{S:methods} describes the methods used by the {\pkg} package to integrate models, \S\ref{S:example} provides a worked example of integrating the two model presented above, \S\ref{S:results} presents several tests demonstrating the performance of message passing between integrated models, and \S\ref{S:discuss} describes a few use cases for {\pkg}, 
summarizes the features and limitations of the {\pkg} package, and outlines areas of ongoing and future development.

@@

{\pkg} is an example of a framework that overcomes these issues by exposing simple and easily accessible interfaces in the languages of the models (See \S\ref{SSS:interface}) that permit messages to be passed between the model processes as they run in parallel (See \S\ref{SS:communication}). As a result, models writers only need knowledge of the language in which their model is written.

@@

{\pkg} is built upon several well-established open-source software packages (see below) along with new tools developed explicitly for {\pkg} including utilities for running and monitoring models written in other languages from Python, dynamically creating and managing communication networks, data type conversions between languages, and asynchronous message passing with variable underlying communication mechanisms. {\pkg} currently support models written in Python, Matlab, C, and C++ with additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}. Support for additional languages is planned for future development (See \S\ref{SS:langsupport}).

@@

Model YAMLs include information about the location of the model source code, the language the model is written in, how the model should be run, and any input or output variables including their data type and units. 
%
%For example, Listings \ref{lst:yml:root} and \ref{lst:yml:shoot} from \A\ref{A:yml} are the model YAMLs for the example root and shoot growth models from \S\ref{S:intro}. These model YAMLs could be used in any integration without being modified because the model inputs and outputs will not change.
%
%\rootyml
%
%\shootyml
%
Integration networks are specified by declaring the connections between models and connections are declared by pairing an output variable from one model with the input variable of another model. The {\pkg} CLI sets up the necessary communication mechanisms to then direct data from one model to the next in the specified pattern. Models can have as many input and/or output variables as is desired and connections between models are specified by references to the input/output variables associated with each model. In addition, input and output variables can also be connected to files.
%
%For example, in the connection YAML for integrating the root and shoot models (Listing \ref{lst:yml:int} from \A\ref{A:yml}), the first two connections link the {\tt root\_input} and {\tt shoot\_input} input channels from the root and shoot model YAML's to input files. The third connection links the {\tt next\_root\_mass} output variable from the root model YAML (Listing \ref{lst:yml:root}) to the {\tt next\_root\_mass} input variable from the shoot model YAML (Listing \ref{lst:yml:shoot}). The forth connection links the {\tt next\_shoot\_mass} to an output file.
%
%\intyml
%
This format allows users the flexibility to create complex integration networks and test models in isolation before running the entire integration network as will be seen in \S\ref{S:example}.

%---------------------------------------------------------------------------------
% MODEL EXECUTION
\subsubsection{Model Execution}\label{SS:model_execution}
%
{\pkg} launches each model in an integration on its own process, allowing models to complete independent operations in parallel and complete tasks more quickly. For example, the root and shoot growth models from \S\ref{S:intro} (described further in \S\ref{S:example}) require 10.28\,s and 10.40\,s respectively to run for 100 time steps in isolation with direct input/output from/to files in their native language. If these two models were manually integrated in serial via the command line, the integration would require a total of 20.68\,s. However, the same integration requires only 16.03\,s using {\pkg}, a speed-up of 1.29. The exact speedup provided by parallel integration using {\pkg} depends on how independent the models are and how equally distributed the work is between the models. If the models are fully independent, the speedup is limited by the time required to execute the slowest model. If the models are dependent on one-another, the models may have to wait to receive messages and the speedup will not be as great. \S\ref{SS:speedup} provides additional information about the speedups achievable through {\pkg} parallel integration and how model structure influences the size of the performance boost.

While every model is executed on a new process, how the model is handled depends on the language it is written in. {\pkg} has a dedicated driver for each of the supported languages with utilities that allow {\pkg} to launch and monitor executables written in that language from Python. Models written in interpreted languages (Python and Matlab) are executed on the command line via the interpreter. In the case of Matlab, where a significant amount of time is required to start the Matlab interpreter (See \S\ref{SS:results_language}), Matlab shared engines are used to execute Matlab models. Matlab shared engines are Matlab instances that other processes can submit Matlab code to for execution. While shared engines also require the same amount of time to start as a standard instance of the Matlab interpreter, they can be started in advance and then reused.

For the compiled languages (C and C++) there are a few options. The user can compile the model themselves, provided they include the source code for the appropriate dependencies at compilation and link against the appropriate {\pkg} header library. {\pkg} provides several command line tools for determining the locations of the necessary libraries and any required compilation/linking flags. Alternatively, users can provide the location of the model source code and let one of several {\pkg} drivers handle the compilation, including linking against the appropriate {\pkg} interface library. {\pkg} also has support for compiling models using Make \citep{Stallman2004} and CMake \citep{Martin2006} for models that already have a Makefile or CMakeLists.txt. To use these tools to compile a model, lines are added to the recipe in order to allow linking against {\pkg}.

@@

The algorithms for a generic model driver and the master thread are provided as Algorithms \ref{alg:model} \& \ref{alg:master} respectively in \A\ref{A:algorithms}.

@@

%---------------------------------------------------------------------------------
% ZEROMQ
\subsubsection{ZeroMQ}\label{SSS:zmq}
%
The preferred communication mechanism used by {\pkg} are ZeroMQ sockets \citep[ZMQ][]{Akgul2013}. ZMQ provides broker-less communication via a number of protocols and patterns with bindings in a wide variety of languages that can be installed on Posix and Windows operating systems. ZMQ was adopted by {\pkg} in order to allow support on Windows and for future target languages (See \S\ref{SS:future}) that could not be accomplished using IPC queues. In addition, while ZMQ allows interprocess communication via IPC, ZMQ also supports protocols for distributed communication via an Internet Protocol (IP) network. While {\pkg} does not currently support using these protocols for distributed integration networks, this is an avenue of future development that has been prepared for.

In addition to using the default ZMQ libraries, {\pkg} also includes supporting routines for allowing received messages to be confirmed. Because ZMQ is broker-less, a socket has no way of knowing if the message it sent was successfully received. This lack of confirmation makes it harder to determine if there is was error somewhere along the network. {\pkg} overcomes this by generating two ZMQ sockets for every connection: one for passing the messages and one for confirming them. Each time a message is received, the receiving socket confirmation thread will send a confirmation including a unique ID for the received message and then wait for a reply to that confirmation. The sending socket confirmation thread will continuously check the confirmation sockets for messages indicating that sent messages were received. Once a confirmation message is received, it will send a reply to the receiving confirmation socket and record the ID of the message that was confirmed. This handshake operation ensures that all messages are accounted for so {\pkg} knows if a message is lost and where it was lost.

@@

%---------------------------------------------------------------------------------
% INTERFACE
\subsubsection{Interface}\label{SSS:interface}
%
{\pkg} provides interface functions/classes for communication that are written in each of the supported languages. Language specific interfaces allow users to program in the language(s) with which they are already familiar.  Each language interface is an implementation of Algorithms \ref{alg:async_send} \& \ref{alg:async_recv} from \A\ref{A:algorithms} that provides users with send and receive calls for passing messages. The Python interface provides communication classes for sending and receiving messages. The Matlab interface provides a simple wrapper class for the Python class, that exposes the appropriate methods and handles conversion between Python and Matlab data types. The C interface provides structures and functions for accessing communication channels and sending/receiving messages. The C++ interface provides classes that wrap the C structures with functions called as methods.

In addition to basic input and output, each interface also provides access to more complex data types and communication patterns that can be found in the {\pkg} documentation. 
%Pseudocode describing the asynchronous communication classes and functions underlying the interfaces is provided by Algorithms \ref{alg:async_send} \& \ref{alg:async_recv} in \A\ref{A:algorithms}.

%---------------------------------------------------------------------------------
% SERIALIZATION
%
\subsection{Transformation}\label{SS:transformation}
%
%\subsubsection{Data Formats}\label{SSS:dataformats}
%
Messages are passed as raw bytes. In order to understand the messages begin passed, parallel processes that communicate must agree upon the format used to do so. Without community standards, different models will often use very different data formats for their input and output. Differences between data formats can include, but are not limited to, type, precision, fields, or units. While some data formats are self-descriptive and include these types of information as metadata, this is not true of all data formats. To combat this, {\pkg} requires models to explicitly specify the format of input and output expected by a model in the model YAML. {\pkg} can then handle a number of conversions between models without prompting as well as serialization/deserialization to the correct type in each of the supported languages. Data formats currently supported by {\pkg} include:
%
\begin{itemize}
	\item Scalars (e.g. integers, decimals)
	\item Arrays
	\item Text-encoded tables (e.g. CSV or tab-delimited)
	\item Pandas data frames \citep{pandas}
	\item 3D geometry structures (PLY \citep{ply} \& OBJ \citep{obj})
\end{itemize}

@@

%---------------------------------------------------------------------------------
%---------------------------------------------------------------------------------
% WORKED EXAMPLE
\section{Worked Example}\label{S:example}
%
In order to illustrate how {\pkg} is intended to be used by model writers, the following walks through the integration of the two examples models from \S\ref{S:intro}. All of the source code and YAML files discussed in this section are available in the {\pkg} GitHub repository and will be included with future releases on PyPI and conda-forge. This example assumes that the user starts with the following existing model source code:
%
\rootsrc
%
\shootsrc
%
The root source code in Listing \ref{lst:src:root} is a standalone C header library containing a single function {\tt calc\_root\_mass} that calculates and returns the root mass at the next time step (a double precision floating point number) from input of the relative root growth rate ({\tt r\_r}, double), the time step ({\tt dt}, double), and the root mass as the current time step ({\tt R\_t}, double).
%
The shoot source code in Listing \ref{lst:src:shoot} is a standalone Python module containing a single function {\tt calc\_shoot\_mass} that calculates and returns the shoot mass at the next time step from input of the relative shoot growth rate ({\tt r\_s}), the time step ({\tt dt}, double), the shoot mass at the current time step ({\tt S\_t}), the root mass at the current time step ({\tt R\_t}), and the root mass at the next time step ({\tt R\_tp1}).
%
In addition to the actual calculation, both models include {\tt sleep} statements (lines 11 \& 18 in Listings \ref{lst:src:root} \& \ref{lst:src:shoot} respectively). These statements are meant to simulate a longer, more involved calculation representative of a more realistic model.

@@

For the example presented above of one model (model A) sending output to another (model B) and then waiting for a response, the speedup offered by {\pkg} through parallelism can be parameterized in terms of the durations of calculations for each model performed in the different phases of the process (See Figure \ref{fig:empty_parallel}). 

@@

%---------------------------------------------------------------------------------
% USE CASES AND LIMITATION
%
\subsection{Advantages \& Limitations}\label{SS:uses}
%
\subsubsection{Advantages}\label{SS:advantages}
%
{\pkg} offers several advantages over manual integrations or other existing tools.
%
\begin{description}
	\item[Dynamic Communication Network] While there are many existing tools for communication that have implementations in multiple languages \citep[some of which {\pkg} uses, e.g.][ZMQ, RabbitMQ, OpenMPI]{ZMQ, RMQ, Gabriel2004}, these tools require that the communication pattern be well established such that it can be hard coded into the model (or model wrapper) itself (e.g. using specific ports or queues). The use of hard coded values greatly restricts the reusability of a model. For every integration a model is used in, the hard coded values for every connection must be updated between the models on either end so that they match. {\pkg} is unique in that it dynamically establishes and manages the networks of communication resources (e.g. queues, channels, sockets) that are required for integrations to run so that models do not need to have these hard coded value and can be reused in multiple integrations, unmodified. This process includes creating new resources, connecting to existing resources, passing the appropriate resource addresses to the model environment, error handling/propagation, and resource cleanup. 
	\item[Hands-off Parallelism \& Asynchronous Communication] In addition to inter-language communication, {\pkg} users benefit from speedups due to parallelism and concurrency without any additional work. From the specified model connections, {\pkg} manages the parallel execution of models and asynchronous message passing under-the-hood so that the user doesn't have to think about it. Existing tools for asynchronous communication between unmanaged parallel tasks \citep[e.g. Jack2,][]{Magoules2018} requires explicit definition of the communication pattern in advance, prohibiting dynamic communication networks as discussed above. Work/data flow tools like Parsl \citep{babuji18} allow dynamic execution of processes along a dependency chain in parallel, but usually require addition code, are limited to passing data as files or data object native to the tool's implementation language, and are ill suited to models with persistent internal states or intermediate communication. {\pkg} is unique in that it offers users asynchronous communication and parallelism that is flexible enough to support any communication pattern, implicitly handles conversion between native data types in the supported languages, and does not require any additional time investment on the part of the user to achieve a speedup.
	\item[No New Language] Unlike DSLs or language specific frameworks, {\pkg} doesn't require a model to be written in one specific language. In addition to expanding the possible number of model combinations, this eases the burden on model writers for whom learning an additional programming language or DSL might be a significant obstacle. {\pkg} is designed such that model writers only need to be familiar with the language that the model is written in to being with. Although {\pkg} is a Python package, users do not even need to be familiar with the Python language and can interact with {\pkg} solely through the command line. While users do need to be able to compose YAML specification files in addition to their model code, these files have a very limited syntax and will be GUI composable in future releases.
\end{description}

\subsubsection{Limitations}\label{SS:limitations}
%
While {\pkg} is a powerful tool for integrating scientific models and opening up new avenues for collaboration, it is not applicable to every scenario. Before diving into model integration, potential users should determine if their use case falls under one of the following scenarios.
%
\begin{description}
	\item[Single Language Integrations] For models written in the same language, it is usually much more computationally efficient to either integrate the models directly (i.e. by calling one from the other) or use a language specific tools for parallel execution and passing data \citep[e.g. Parsl in Python][]{babuji18} than to use {\pkg}. This performance advantage will be particularly pronounced for models that finish quickly or tightly coupled integrations where models make a large number of exchanges in rapid succession (e.g. a set of coupled ODEs). Although the constituent models will still benefit from integration via {\pkg} in regards to reusability, {\pkg} may not be the best choice in such cases if execution time is a concern. Future work on automated creation of wrapper files for function-like (\S\ref{SS:control}) and ODE (\S\ref{SS:ODE}) models may allow {\pkg} to handle this automatically, replacing managed communication with in-language variable passing for models that are written in the same language and co-located on the same machine.
	\item[Fast \& Tightly Coupled Models] For models involving quick calculations or for tightly couple sets of models such as might occur for two models representing coupled ODEs, the computational cost of rapid communication via {\pkg} can quickly surpass the cost of the calculations themselves causing the {\pkg} to be computationally inefficient in comparison with a manual integration. For example, if the sleep statements were removed from the example root and shoot models from \S\ref{S:example}, the calculations become trivial and take significantly less time than the communication between the models. Without the sleep statements, the root and shoot models run for 100 time steps in 0.007\,s and 0.095\,s respectively while the integration takes 7.437\,s. The speedup offered by {\pkg} is then 0.0137, which is actually a \emph{significant} slow down. For such models, large parameter searches or integrations with many time steps can be computationally prohibitive with {\pkg}.
	\item[Unsupported Languages] While {\pkg} currently supports models written in Python, Matlab, C, \& C++, these languages are not an exhaustive list of those languages in which models are written. There are plans to expand support for {\pkg} to other languages (see \S\ref{SS:langsupport}), but in the meantime {\pkg} will not be particularly useful for models that are not written in one of the languages already supported. However, if a language can be called from one of the supported languages or vice versa \citep[e.g. f2py or SWIG,][]{Peterson2009,Beazley2003}, it may be possible to either write a wrapper or make calls to an interface in a supported language.
\end{description}

%---------------------------------------------------------------------------------
% FUTURE WORK
\subsection{Ongoing/Future Improvements}\label{SS:future}
%
{\pkg} is being actively developed to expand the number of models that can be used in integration networks and the complexity of integration networks that can be executed. Several improvements are already in progress/planned for {\pkg}.

% LANGUAGE SUPPORT
\subsubsection{Language Support}\label{SS:langsupport}
%
The biggest barrier to running new models is language support as plant models are written in many different languages. While the current language support covers many of the most popular languages among plant biologists, additional languages must be added to unlock the full set of potential integrations. Informal surveys of the plant modeling community during recent conferences and workshops have helped identify the core languages being used by model developers that would have the larges impact on the number of models accessible via {\pkg}. Based on these results, we plan to expand {\pkg} to support models written in R, Fortran, Java, and the SBML \citep{Hucka2003} DSL during the next major development push with support for additional models in subsequent releases as requested.

@@

% DISTRIBUTED SYSTEMS
\subsubsection{Distributed Systems}\label{SS:distributed}
%
High performance computing (HPC) and cloud compute resources are powerful tools in the current computing ecosystem that could be used for running complex integration networks. To this end, we plan to expand {\pkg} support for running integration networks on distributed compute resources. {\pkg} already uses communication tools that can be adapted for use in a distributed pattern \citep[ZMQ, RMQ][]{ZMQ, RMQ} and we will leverage tools like the libsubmit package from the Parsl project \citep{babuji18} for automating the submission process to HPC and compute resources. This work will also enable {\pkg} to adaptively spawning additional model processes on elastic compute resources when a model becomes a bottle neck to the integration process. This feature will be particularly useful for integrations simulating large ensembles of biological units (e.g. leaves on a plant or plants in a field) and will enable scientists to explore emergent properties of these systems.

In addition, we will also add tools to {\pkg} for running models as a service and using RMQ to permit access to these models within integration networks. This feature will be useful for scientists who are unable or unwilling to share the source code for their model. Instead, these models can be hosted on dedicated compute resources managed by the model's home institution and collaborators interested in using the model in an integration can connect to it remotely, submitting requests and receiving results as needed.

% CONTROL FLOW
\subsubsection{Wrapper Automation \& Control Flow}\label{SS:control}
%
Currently, modelers must explicitly decide how a model will be used when writing the model wrapper. These decisions can include things like which input and output variables are static versus variable, if there are sets of variables that should be updated at the same time, and how often output is generated, resulting in a hard coded use pattern. If someone wishing to use the model in another integration requires a different use patter, they will need to write their own wrapper. However, for models that have a functional form, the process of writing the model wrapper can be automated and done dynamically for each integration. We plan on adding options to {\pkg} for dynamically generating model wrappers based a user provided function call and list of static/variable input channels. This feature will improve the reusability of model code such as during parameter studies and decrease the amount of work required for integrating models that are written or can be exposed in a functional form.
